---
title: "Ownership map scraps"
author: "Brenna Robeson"
date: "July 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## RUN 1 SCRAPS

### Small-scale attempt with Worc. Academy Census Block Group

Read in all data
```{r}
# read in MA census block groups
macbg <- read_sf("C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/CENSUS2010BLOCKGROUPS_POLY.shp") 

# project maejpop to EPSG:32618, WSG84 UTM Zone 18N
macbg <- st_transform(x = macbg, crs = st_crs("+proj=utm +zone=18 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))

# filter out only Worcester block groups
cbg <- macbg %>% st_as_sf() %>% filter(GEOID10 %in% 250277301001:250277331022)

# read in property parcels (from Worcester 2019 geodatabase)
parcels <- read_sf("C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/Basemap_Parcels_Polygons_19.shp")
head(parcels)

# project parcels to EPSG:32618, WSG84 UTM Zone 18N
parcels <- st_transform(x = parcels, crs = st_crs("+proj=utm +zone=18 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")) %>% st_as_sf()

# ### PARCELS TEST
# parcels19 <- st_read(dsn = "C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/Worcester2019.gdb", layer = "Basemap_Parcels_Polygons")
# parcels19 <- st_transform(x = parcels19, crs = st_crs("+proj=utm +zone=18 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")) %>% 
#   as_tibble %>% st_as_sf()
# ###

# read in property parcel ownership database (from ESRI geodatabase) as tibble
parcelsdb <- st_read(dsn = "C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/Worcester2019.gdb", layer = "BASEMAP_PARCELS_DATABASE", as_tibble = TRUE)
head(parcelsdb)
```

Prepare data associated with WA census block group for geocoding
```{r}
# clip parcels data with cbg geometry for WA
wacbg <- cbg %>% filter(LOGSF1 == 171787) %>% st_as_sf()
parcelsinter <- st_intersection(x = wacbg, y = parcels)
#parcelswa19 <- st_intersection(x = wacbg, y = parcels19)

# # test plot wacbg
# par(mar = rep(0, 4))
# # #plot(st_geometry(parcels), col = "grey")
# plot(st_geometry(parcelswa), col = rainbow(n = nrow(parcelswa)))

# join property parcels with their ownership data using a left_join using database field "GIS_MBL" and shapefile field "MAP_PAR_ID" and create single address field

# 6/23/19: need to find parcels with multiple database records to explain discrepency between 184 parcels and 199 observations in this new variable after left_join
# HOW TO PICK OUT WHICH PARCELS HAVE MULTIPLE RECORDS (IF THIS IS EVEN PROBLEM) AND WHAT TO DO ABOUT IT...
parcelswa <- left_join(parcelsinter %>% select(GEOID10, POP100_RE, SHAPE_AREA, MAP_PAR_ID) %>% as.data.frame, parcelsdb %>% select(CLASS, OWNER, RODYEAR, MAIL_STRET, MAIL_CITY, MAIL_STATE, MAIL_ZIP5, VALUE, ZONE, ZONE_NAME, BLDDESC, HOUSE_NO, STREET_PRE, STREET, LND_VALU, BLD_VALU, MAIL_STR2, GIS_MBL) %>% as.data.frame, by = c("MAP_PAR_ID" = "GIS_MBL")) %>% 
  as_tibble %>% st_as_sf() #%>% 
  #unite(MAIL_FULL, )

#############
# WHY ARE THERE MORE RECORDS(199) THAN PARCELS(184)???

# export object out of R to ArcMap to investigate duplicates
st_write(parcelswa, "C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/ownership_test/parcelswadb.shp")

# how many records if group_by?
parcelswa_g <- parcelswa %>% group_by(MAP_PAR_ID) %>% summarize(OWNERCNT = __??__(MAP_PAR_ID))

parcelswa_g <- parcelswa %>% tally(.$MAP_PAR_ID, sort = FALSE, name = "OWNERCNT")

#############

# unite(EIJ_GRP, WHITE, AFFLUENT, ENGLISHPROF, sep = "", remove = FALSE) %>% 
#   mutate(EIJ_GRP = gsub("NA", "", EIJ_GRP)) 

#length(unique(bmorehoods@data$[id]))

# aggregate (with unite?) all the owner address fields

# filter out those to be geocoded (aka parcels with ownership in worcester and outside of the block group?)

# (if any llcs or pobox numbers, look up these addresses?)

# convert these selected rows into a data frame
```


Geocode ownership information
```{r}
# use ggmap::mutate_geocode() to geocode ownership addresses

# NEED TO DO MORE RESEARCH
# OPT 1: just count how many properties registers per census block group using st_join?? and see where the ownership locations fall... is it in EIJ or EJ areas?
# OPT 2: figure out how to get lines between parcels and ownership parcels?...
```





## RUN 2 SCRAPS


### playing with previous scraps

Read in and recreate Environmental Injustice Population data
```{r, warning = FALSE, message = FALSE, error = TRUE}
# data source: see eijpop vignette

# read in ej pop data
maejpop <- read_sf("C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/EJ_POLY.shp")

# project maejpop to EPSG:32618, WSG84 UTM Zone 18N
maejpop <- st_transform(x = maejpop, crs = st_crs("+proj=utm +zone=18 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))

# filter out Worcester-only data and view first few rows
ejpop <- maejpop %>% st_as_sf() %>% filter(TOWN == "WORCESTER")
head(ejpop)

# read in MA census block groups
macbg <- read_sf("C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/CENSUS2010BLOCKGROUPS_POLY.shp") 

# project maejpop to EPSG:32618, WSG84 UTM Zone 18N
macbg <- st_transform(x = macbg, crs = st_crs("+proj=utm +zone=18 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))

# filter out only Worcester block groups
cbg <- macbg %>% st_as_sf() %>% filter(GEOID10 %in% 250277301001:250277331022)
head(cbg)

# join ejpop data to cbg data
ejjoincol <- ejpop %>% select(LOGSF1, MINORITY, INCOME, ENGLISH, EJ_CRITERI) 
cbg_ejpop <- left_join(cbg %>% select(GEOID10, LOGSF1) %>% as_tibble, 
                       ejjoincol %>% as.data.frame() %>% select(-geometry), 
                       by = "LOGSF1") %>% st_as_sf()

# # test results of join
# ejjoincol %>% filter(LOGSF1 == 171758)
# cbg %>% filter(LOGSF1 == 171758)
# cbg_ejpop %>% filter(LOGSF1 == 171758)
# 
# cbg_ejpop %>% st_as_sf() %>% select(MINORITY) %>% plot()
# cbg_ejpop %>% filter(!is.na(MINORITY))

# create new columns WHITE, AFFLUENT, ENGLISHPROF, EIJ_CRITERI in cbg_ejpop
cbg_ejpop <- cbg_ejpop %>% 
  mutate(WHITE = ifelse(is.na(MINORITY), "W", NA)) %>% 
  mutate(AFFLUENT = ifelse(is.na(INCOME), "A", NA)) %>% 
  mutate(ENGLISHPROF = ifelse(is.na(ENGLISH), "P", NA)) %>% 
  unite(EIJ_GRP, WHITE, AFFLUENT, ENGLISHPROF, sep = "", remove = FALSE) %>% 
  mutate(EIJ_GRP = gsub("NA", "", EIJ_GRP)) 

# create eij pop which includes only eij populations
eijpop <- cbg_ejpop %>% filter(EIJ_GRP != "")
```


```{r, warning = FALSE, message = FALSE, fig.width=6, fig.height=4, fig.align='center'}
# read in parcel shape file
parcels <- read_sf("C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/Basemap_Parcels_Polygons.shp")
head(parcels)

# project parcels to EPSG:32618, WSG84 UTM Zone 18N
parcels <- st_transform(x = parcels, crs = st_crs("+proj=utm +zone=18 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")) %>% st_as_sf()

# intersect parcel data wtih object from ejpop census block group to show subset of parcel data
testarea <- ejpop %>% filter(LOGSF1 == 171787) %>% st_as_sf()
  
  
  #slice(1) %>% st_as_sf

# par(mar = rep(0, 4))
# plot(st_geometry(parcels), col = "grey")
# plot(st_geometry(testarea), col = "purple", add = TRUE)
parcels_testarea <- st_intersection(x = testarea, y = parcels)
par(mar = rep(0, 4))
#plot(st_geometry(parcels), col = "grey")
parcels_subset <- plot(st_geometry(parcels_testarea), col = rainbow(n = nrow(parcels_testarea)))

```





## RUN 3 SCRAPS



```{r}
### TEST - GEOCODING #############################
set.seed(1)
fdp_test <- fdparcels %>% sample_n(., size = 50, replace = FALSE)

# create full address column
fdp_test <- unite(fdp_test %>% as.data.frame(), ADDRESS_FULL, MAIL_STRET, MAIL_CITY, MAIL_STATE, MAIL_ZIP5, sep = ", ") %>% as_tibble()
# create full address column that references MAIL_STR2 when c/o present
fdp_test2 <-  fdp_test %>% mutate(ADDRESS_FULL = ifelse(str_detect(MAIL_STRET, "C/O"), unite(fdp_test %>% as.data.frame(), ADDRESS_FULL, MAIL_STR2, MAIL_CITY, MAIL_STATE, MAIL_ZIP5, sep = ", ", remove = FALSE), unite(fdp_test %>% as.data.frame(), ADDRESS_FULL, MAIL_STRET, MAIL_CITY, MAIL_STATE, MAIL_ZIP5, sep = ", ", remove = FALSE)))
# if there are an overwhelming amount of addresses with c/o, use ifelse
# fdp_test <- fdp_test %>% mutate(ADDRESS_FULL = ifelse(str_detect(MAIL_STRET, ), yes, no),
#                                 ADDRESS_FULL = ifelse(test, yes, no))
#   unite(. %>% as.data.frame(), ADDRESS_FULL, )



# use address column to create new column of coordinates for ownership addresses of parcels (convert to data.frame)
# google aip AIzaSyBjBeWf5pvV37byqx82Vlo854-vfie3v_0
fdp_test_gc <- fdp_test %>% as.data.frame() %>% 
  mutate_geocode(., ADDRESS_FULL) %>% 
  as_tibble()
head(fdp_test_gc)


### TEST - PLOTTING PARCEL COORDS ################
# jk because they're from the data being randomly projected

################################################
```


### RUN 4?? 7/18/19 SCRAPS

Geocode owner addresses for each parcel
```{r}
### REGISTER GOOGLE API KEY #####################
register_google(key = "AIzaSyBjBeWf5pvV37byqx82Vlo854-vfie3v_0", write = TRUE)
################################################

# # create more comprehensive address column to address "c/o"s in MAIL_STRET
# fdparcels <- fdparcels %>% mutate(MAIL_STR_CMB = MAIL_STR) %>% # how to identify c/os??? grepl/grep selection isn't working

# assess how many might be po box numbers...
# what does ggmap do with invalid/unlocated addresses?

# # concatenate different address pieces into a full address for geocoding
# fdparcels_add <- unite(fdparcels %>% as.data.frame(), ADDRESS_FULL, MAIL_STRET, MAIL_CITY, MAIL_STATE, MAIL_ZIP5, sep = ", ", remove = FALSE) %>% as_tibble()
# 
# # create seperate object to geocode owner addresses
# ownership <- fdparcels_add %>% select(MAP_PAR_ID, ADDRESS_FULL)
# 
# # use address column to geocode ownership locations
# ownership_gc <- ownership %>% as.data.frame() %>% 
#   mutate_geocode(., ADDRESS_FULL) %>% 
#   as_tibble()
# head(ownership_gc)

# # save results of geocoding to .csv 
# st_write(ownership_gc, "C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/ownership_gccoord.csv")

# read in csv with geocoded ownership addresses
ownership_gc <- read_sf("C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/fdej_ownership_gc.csv")

# view geocoded ownership coordinate locations
#filter(!is.na(lon))
ownership_gc <- ownership_gc %>% select(MAP_PAR_ID, lon, lat) %>% filter(!lon == "") %>% st_as_sf(., coords = c("lat", "lon"), crs = 4326)
mapview(ownership_gc)
#st_crs(ownership_gc) <- "+proj=utm +zone=18 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

# %>% # does setting crs actually project data?
#   st_transform(x = ., crs = st_crs("+proj=utm +zone=18 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))

#st_crs(ownership_gc_up) <- st_transform(ownership_gc, 4326)
fdeijpop_up <- fdeijpop %>% st_transform(crs = st_crs(ownership_gc))

sfdat <- list("fdeijpop_up" = fdeijpop_up,  "ownership_gc" = ownership_gc)
sapply(sfdat, function(x) st_bbox(x))

par(mar = c(0, 0, 0, 0))
plot(st_geometry(fdeijpop_up), col = "grey")
plot(st_geometry(ownership_gc), col = "blue", pch = 20, add = TRUE)


# # sf back to tibble
# ownership_gc_p <- bind_cols(ownership_gc %>% as_tibble %>% select(-geometry), 
#           st_coordinates(ownership_gc) %>% as_tibble)
# 
# # try plotting in google maps with ggmap
# p <- ggmap(get_googlemap(center = c(lon = 41.9, lat = -71.0),
#                     zoom = 11, scale = 2,
#                     maptype ='terrain',
#                     color = 'color'))
# p + geom_point(aes(x = X, y = Y,  colour = "blue"), data = ownership_gc_p, size = 0.5) + 
#   theme(legend.position="bottom")

# # export ownership_gc_coord as .shp
# st_write(ownership_gc, "C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/fdej_ownership_gc.shp")

# what are the coordinates of fdejpop centroids versus ownership coords?
fdej_cent <- st_centroid(fdejpop)

```


## scraps from 7.23.19 edits

```{r}
# trying to properly map geocoded ownership addresses
# proj4string(gc) <- CRS("+proj=longlat +datum=WGS84")
# gc <- spTransform(gc, CRS(proj4string(country)))
# 
# # checking bbox
# sfdat <- list("fdeijpop_up" = fdeijpop_up,  "ownership_gc" = ownership_gc)
# sapply(sfdat, function(x) st_bbox(x))



# # sf back to tibble
# ownership_gc_p <- bind_cols(ownership_gc %>% as_tibble %>% select(-geometry), 
#           st_coordinates(ownership_gc) %>% as_tibble)
# 
# # try plotting in google maps with ggmap
# p <- ggmap(get_googlemap(center = c(lon = 41.9, lat = -71.0),
#                     zoom = 11, scale = 2,
#                     maptype ='terrain',
#                     color = 'color'))
# p + geom_point(aes(x = X, y = Y,  colour = "blue"), data = ownership_gc_p, size = 0.5) + 
#   theme(legend.position="bottom")



# what are the coordinates of fdejpop centroids versus ownership coords?
#fdej_cent <- st_centroid(fdejpop)

```




magic of taylor
get points for parcels within EJ
geocode addresses of owners -> get points for this as well as output
write some magic formula (or maybe read about spatial network analysis) - if the point of a property's owner falls in EIJ, connect the two with a line,otherwise disregard?
  maybe connect everything at first if it's not too overwhelming for system to see signifigance of relationship (but can't assess sig without stats, so mabye skip that...)
  orrrr - aggregate hits per census block group, then choropleth map of totals

### hellooo it's saturday 7.27.19

```{r}
### attempt 1 ##############################
# need title and legend
# ownership_eij2ej <- ggplot() +
#   geom_sf(data = worcester, aes(fill = "gray90", col = "gray")) +
#   geom_sf(data = fdeijpop, aes(fill = "darkgoldenrod1", col = "gray60")) +
#   geom_sf(data = fdejpop, aes(fill = "orchid", col = "gray60")) +
#   geom_sf(data = fdparcels_wown, aes(col = "darkslateblue"), pch = 20) +
#   geom_sf(data = own_fdeij, aes(col = "orangered2")) +
#   ggtitle("Relational Property Ownership ") +
#   coord_sf(datum = NA)
# ownership_eij2ej

### attempt 2 ##############################

```
