---
title: 'Perspective 3: Food Desert Rent Extraction'
author: "Brenna Robeson"
date: "May 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(igraph)
library(dplyr)
library(sf)
library(tidyverse)
library(ggmap) # add citation??...
library(mapview) # get rid of any references when you knit?
library(ggplot2)
```


## Objective

Examine the ownership arrangements of property parcels found within food deserts to see if and how property ownership acts as extractive and is thus destabilizing to community autonomy over their neighborhood dynamics and food systems

Eventually will include a choropleth map of rentership rates by geographical unit used in analysis


## Mapping Process


Read in data
```{r}
# read in EJ Populations
maejpop <- read_sf("C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/EJ_POLY.shp") 

# project maejpop to EPSG:32618, WSG84 UTM Zone 18N
maejpop <- st_transform(x = maejpop, crs = st_crs("+proj=utm +zone=18 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))

# filter out worcester
ejpop <- maejpop %>% st_as_sf() %>% filter(TOWN == "WORCESTER")

# read in EIJ Populations
eijpop <- read_sf("C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/eijpop.shp")

# project EIJ populations
eijpop <- st_transform(x = eijpop, crs = st_crs("+proj=utm +zone=18 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))

# fix EIJ pop column names 
# HOW TO FIX THIS UPSTREAM? (like in eijpop.Rmd?)
eijpop <- eijpop %>% rename(AREA_SQFT = AREA_SQ, MINORITY = MINORIT, EJ_CRITERI = EJ_CRIT, AFFLUENT = AFFLUEN, ENGLISHPROF = ENGLISHP)

# read in Worcester Census Tracts
ct <- read_sf("C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/Census_Tracts_2010.shp") 

# project wct to EPSG:32618, WSG84 UTM Zone 18N
ct <- st_transform(x = ct, crs = st_crs("+proj=utm +zone=18 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))

# read in food desert csv
fd <- read_sf("C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/DesAtlas_2015_WorcSel.csv")

# join fd .csv to census tract geometries
fdct <- left_join(ct %>% as.data.frame, fd, by = c("GEOID10" = "CensusTract")) %>% 
  as_tibble %>% st_as_sf()
```

Select census block groups to use for analysis
*THINK ABOUT THEORETICAL UNDERPINNING FOR SELECTING THESE TWO GROUPS - WHAT DOES IT CONTRIBUTE TO YOUR RESEARCH QUESTION? OTHER POP YOU SHOULD BE THINKING ABOUT?... theoretical underpinning? wouldn't it be crazy if these areas of disinvestment where marginalization is concentrated were connected to these areas where privelage is super concentrated?? maybe it wouldn't actually because racial capitalism*

```{r}
# select EJ Populations (MIE) which fall in food deserts
fdejpop <- st_intersection(ejpop %>% filter(EJ_CRITERI == "MIE"), fdct %>% 
                             filter(LILATracts_1And10 == "1" | LILATracts_halfAnd10 == "1")) %>%
  st_buffer(., dist = 0, endCapStyle = "ROUND", joinStyle = "ROUND") %>% 
  mutate(INT_AREA = st_area(.) %>% measurements::conv_unit(., "m2", "ft2") %>% as.numeric) %>% 
  st_as_sf() %>% 
  filter(near(AREA_SQFT, INT_AREA, tol = 100000))

# write fdej to shapefile
st_write(fdejpop, "C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/fdejpop.shp")
  
# select EIJ Populations (AWP) which fall outside of food deserts
fdeijpop <- st_intersection(eijpop %>% filter(EIJ_GRP == "WAP"), fdct %>% filter(LILATracts_1And10 == "0" | LILATracts_halfAnd10 == "0")) %>%
  st_buffer(., dist = 0, endCapStyle = "ROUND", joinStyle = "ROUND") %>% 
  mutate(INT_AREA = st_area(.) %>% measurements::conv_unit(., "m2", "ft2") %>% as.numeric) %>% 
  st_as_sf() %>% 
  filter(near(AREA_SQFT, INT_AREA, tol = 100000))

# # write fdeij to shapefile
# st_write(fdeijpop, "C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/fdeijpop.shp")
```


Prepare Parcels Data
```{r}
# read in property parcels (from Worcester 2019 geodatabase)
parcels <- read_sf("C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/Basemap_Parcels_Polygons_19.shp") 
head(parcels)

# project parcels to EPSG:32618, WSG84 UTM Zone 18N
parcels <- st_transform(x = parcels, crs = st_crs("+proj=utm +zone=18 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")) %>% st_as_sf()

# clip parcels data to only include area of food desert ejpops
parcelsinterfd <- st_intersection(x = fdejpop %>% select(AREA_SQFT, POP100_RE, MED_HH_INC, PCT_MIN, ENGLISH, INCOME, MINORITY, EJ_CRITERI, LILATracts_1And10, LILATracts_halfAnd10, LILATracts_1And20, LILATracts_Vehicle), y = parcels)

# read in property parcel ownership database (from ESRI geodatabase) as tibble
parcelsdb <- st_read(dsn = "C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/Worcester2019.gdb", layer = "BASEMAP_PARCELS_DATABASE", as_tibble = TRUE)
head(parcelsdb)

# join parcels database to fdejpop parcels shapefile (with respective fields "GIS_MBL" and "MAP_PAR_ID")
fdparcels <- left_join(parcelsinterfd %>% as.data.frame, parcelsdb %>% #select(CLASS, OWNER, RODYEAR, MAIL_STRET, MAIL_CITY, MAIL_STATE, MAIL_ZIP5, VALUE, ZONE, ZONE_NAME, BLDDESC, HOUSE_NO, STREET_PRE, STREET, LND_VALU, BLD_VALU, MAIL_STR2, GIS_MBL) %>% 
                         as.data.frame, by = c("MAP_PAR_ID" = "GIS_MBL")) %>% 
  as_tibble %>% st_as_sf() # do I need these??

# convert parcel geometries to points

# filter out residential properties only based on ZONE_NAME
res <- c("RESIDENTIAL", "RESIDENT")
fdparcels <- fdparcels %>% 
  filter(., str_detect(ZONE_NAME, paste(res, collapse="|")))
```


Geocode owner addresses for each parcel
```{r}
### REGISTER GOOGLE API KEY #####################
# register_google(key = "AIzaSyBjBeWf5pvV37byqx82Vlo854-vfie3v_0", write = TRUE)
################################################

# # create more comprehensive address column to address "c/o"s in MAIL_STRET
# fdparcels <- fdparcels %>% mutate(MAIL_STR_CMB = MAIL_STR) %>% # how to identify c/os??? grepl/grep selection isn't working

# assess how many might be po box numbers...
# what does ggmap do with invalid/unlocated addresses?

# # concatenate different address pieces into a full address for geocoding
# fdparcels_add <- unite(fdparcels %>% as.data.frame(), ADDRESS_FULL, MAIL_STRET, MAIL_CITY, MAIL_STATE, MAIL_ZIP5, sep = ", ", remove = FALSE) %>% as_tibble()
# 
# # create seperate object to geocode owner addresses
# ownership <- fdparcels_add %>% select(MAP_PAR_ID, ADDRESS_FULL)
# 
# # use address column to geocode ownership locations
# ownership_gc <- ownership %>% as.data.frame() %>% 
#   mutate_geocode(., ADDRESS_FULL) %>% 
#   as_tibble()
# head(ownership_gc)

# # save results of geocoding to .csv 
# st_write(ownership_gc, "C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/ownership_gccoord.csv")

# read in csv with geocoded ownership addresses
ownership_gc <- read_sf("C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/fdej_ownership_gc.csv")

# clean geocoded ownership coordinate locations
ownership_gc <- ownership_gc %>% #select(MAP_PAR_ID, lon, lat) %>%  NEED TO DELETE THIS TO INCLUDE INFO ABOUT PROPERTIES AND THEIR VALUE ETC - 1. MAKE SURE IT DOESN'T MESS WITH DOWNSTREAM STEPS 2. FILTER OUT EIJ AND FOOD DESERT COLUMNS
  rename(lon = lat, lat = lon)

# convert csv with ownership locations to sf object
ownership_gc <- ownership_gc %>% filter(!lon =="") %>% 
  st_as_sf(., coords = c("lat", "lon"), crs = 4326) %>% 
  st_transform(x = ., crs = st_crs("+proj=utm +zone=18 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))
 
# explore location of ownership points relative to fdej and fdeij
par(mar = c(0, 0, 0, 0))
plot(st_geometry(fdeijpop), col = "grey")
plot(st_geometry(fdejpop), col = "orange", add = TRUE)
plot(st_geometry(ownership_gc), col = "blue", pch = 1, add = TRUE)
#mapview(ownership_gc)

# # export ownership_gc as .shp
# st_write(ownership_gc, "C:/Users/Brenna Robeson/OneDrive/Masters_Research_Project/Empirics/GIS_Analysis/extdata/fdej_parcel_ownership.shp")

# Review literature review and design questions to ask of the data...

# Which owners are located within the food desert MEI areas?
own_fdej <- st_intersection(x = fdejpop %>% select(GEOID10, geometry), y = ownership_gc)
#mapview(own_fdej)

# Which owners are located within the non-food desert WAP areas?
own_fdeij <- st_intersection(x = fdeijpop %>% select(GEOID10, geometry), y = ownership_gc)
#mapview(own_fdeij)

# playing
# par(mar = c(0, 0, 0, 0))
# plot(st_geometry(fdejpop), col = "grey")
# plot(st_geometry(own_fdeij), col = "blue", pch = 20, add = TRUE)

# join parceldb data back 
```

Preparing data for plotting - functional map...
```{r}
# load in towns data to extract worcester boundary
worcester <- read_sf("C:/Users/Brenna Robeson/Documents/MassINC_Data/town_geo/gcboundaries.shp") %>% 
  st_transform(., crs = st_crs("+proj=utm +zone=18 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")) %>% filter(TOWN == 'WORCESTER')

# convert fdparcels into centroids
fdparcels_wown <- inner_join(x = fdparcels %>% as.data.frame, y = own_fdeij %>% as.data.frame, by = "MAP_PAR_ID") %>% 
  st_as_sf %>% 
  st_centroid()
 
# plot parcels 
ownership_eij2ej <- ggplot() +
  geom_sf(data = worcester, fill = "gray", col = NA, show.legend = FALSE) +
  geom_sf(data = fdeijpop, aes(fill = "Non-Food Desert EIJ WAP Pop")) +
  geom_sf(data = fdejpop, aes(fill = "Food Desert EJ MIE Pop")) +
  geom_sf(data = fdparcels_wown, aes(colour = "Property Parcel"), pch = 20, show.legend = "point") +
  geom_sf(data = own_fdeij, aes(colour = "Owner Location"), show.legend = "point") +
  scale_fill_manual(name = "", values = c("Non-Food Desert EIJ WAP Pop" = "darkgoldenrod1", "Food Desert EJ MIE Pop" = "orchid"), 
                    guide = guide_legend(override.aes = list(linetype = c("blank", "blank"), shape = c(NA, NA)))) +
  scale_colour_manual(name = "", values = c("Property Parcel" = "darkslateblue", "Owner Location" = "orangered2"),
                      guide = guide_legend(override.aes = list(linetype = c("blank", "blank"), shape = c(16, 20)))) +
  theme(legend.key = element_rect(fill = "white")) +
  ggtitle("Relational Property Ownership Patterns \nBreaching Food Desert Boundaries in Worcester, MA") +
  theme_void() + # Empty theme without axis lines and texts
  theme(
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.box.background = element_rect(fill = "transparent", colour = NA)
  ) +
  coord_sf(datum = NA)
###############################################
# need to lighten census bg boundaries
# fix legend lables
# get rid of "color" and "fill" sections in legend to have everything grouped together - https://stackoverflow.com/questions/12410908/combine-legends-for-color-and-shape-into-a-single-legend
# make points each same size? or more clearly distinct shape?
# title centered?
ownership_eij2ej

# export map to .png
ggplot2::ggsave(ownership_eij2ej, filename = "C:/Users/Brenna Robeson/OneDrive/GEOG346/worcfoodaccess/figures/ownership_plot.png", 
       width = 6, height = 5, units = "in", dpi = 300)

```




## Limitations


## Conclusions
no one aspect or discourse concerning unjust food access holds all the answers. however the food desert narrative is insufficient and problematic in ways that specifically perpetuate structural injustices that have long been recognized by many scholars, residents, and activists as the root drivers of many unjust patterns.

the food desert narrative is lucrative because white, professional people invisibilize themselves in the story of the problem and in the maps of the problem. we are intentionally asserting these maps here because white rich people's role in the creation and perpetuation of the probelem of unjust food access demands conceptualization and exploration. Thus, this set of maps begins to open up space to question who and what is invisibilized and intentionally excluded from traditional maps and what directions future mappings of food access can take to inform policy that addresses the issue through a power-dynamic rather than deficit appraoch.

*THIS IS YOUR ARGUMENT - LOOK AT POWER DYNAMICS, NOT DEFICITS*



