---
title: "Playing with tidycensus package"
author: "Brenna Robeson"
date: "April 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)      # For plotting
library(tidycensus)   # For downloading Census data
library(tmap)         # For creating tmap
library(tmaptools)    # For reading and processing spatial data related to tmap
library(dplyr)        # For data wrangling
library(sf)        
```

### Getting set up with API Census Key

```{r}
census_api_key("32fb923db5a2b235a2824bfcc75ef908ce7aa293", install = TRUE, overwrite = TRUE)
readRenviron("~/.Renviron")
```

### Tutorial from zevross.com blog on healthcare

Tutorial on health insurance coverage for different counties

#### downloading data with get_acs()
```{r}
CTdat12 <- get_acs("county", table = "B27001", year = 2012, 
  output = "tidy", state = 09, geometry = FALSE) %>%
  rename(`2012` = estimate) %>%
  select(-NAME, -moe) 

CTdat16 <- get_acs("county", table = "B27001", year = 2016, 
  output = "tidy", state = 09, geometry = TRUE) %>%
  rename(`2016` = estimate) %>%
  select(-moe)
```

#### Process data

```{r}
dat <- left_join(CTdat16, CTdat12, by = c("GEOID", "variable"))
st_geometry(dat) <- NULL # This drops the geometry and leaves a table
head(dat)
```
Assign heatlh insurance categories
```{r}
dat <- mutate(dat,
  cat = case_when(
    variable %in% paste0("B27001_0",
      c("09","12","37","40")) ~ "pop1834",
    variable %in% paste0("B27001_0",
      c("11","14","39","42")) ~ "pop1834ni")) %>%
  filter(!is.na(cat))
```

summarize the data by county-year-category
```{r}

# Create long version
dat <- tidyr::gather(dat, year, estimate, c(`2012`, `2016`))

# Group the data by our new categories and sum
dat <- group_by(dat, GEOID, NAME, year, cat) %>%
  summarize(estimate = sum(estimate)) %>%
  ungroup() %>%
  tidyr::spread(cat, estimate)

head(dat)
```

calculate final percentage estimates
```{r}
dat <- mutate(dat, est = (pop1834ni/pop1834) * 100) %>%
  select(-c(pop1834, pop1834ni)) %>%
  tidyr::spread(year, est) %>%
  mutate(diff = `2016`-`2012`)

head(dat)
```

#### Visualize

county distributions and median pct uninsured by year using `facet_wrap`
```{r}
datlong <- select(dat, -diff) %>%
  tidyr::gather(year, estimate, c(`2012`, `2016`)) %>%
  group_by(year) %>%
  mutate(med = round(median(estimate, na.rm = TRUE), 1))

ggplot(datlong, aes(estimate)) +
  geom_histogram(fill = "firebrick2", 
    color = "white", bins = 60) +
  xlab("Uninsured adults ages 18-34 by county (%)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~year, ncol = 1) +
  geom_vline(aes(xintercept = med,
    group = year), lty = "dashed") +
  geom_text(aes(label = paste("Median = ", med), x = med, y = 55))
```

(ggplot configurations not relevant because only using CT data)

find counties with the greatest change in pct uninsured with `dplyr::top_n`

```{r}
d2 <- top_n(dat, 2, diff) %>%
  mutate(type = "Insured population decreased",
    difftemp = diff)

i2 <- top_n(dat, -2, diff) %>%
  mutate(type = "Insured population increased",
    difftemp = abs(diff))

id2 <- bind_rows(list(i2, d2)) %>%
  arrange(desc(difftemp))
```

```{r}
ggplot(id2) + 
  geom_col(aes(x = forcats::fct_reorder(NAME, difftemp), 
    y = difftemp, fill = type)) +
  coord_flip() +
    scale_fill_manual(values = c("firebrick2", "cyan4")) +
  theme(plot.title = element_text(hjust = 0.5),
    legend.position = "bottom",
    legend.title = element_blank()) +
  ggtitle("Counties with the greatest change (+/-) in
    insured population, ages 18-34, 2012-2016") +
  ylab("Difference in % insured (2016 - 2012)") +
  xlab("")
```

Create a geographic file with data (to be fed into `tmap`)


```{r}
# dat16 is our original geographic object and dat is the tabular data
shp <- CTdat16 %>%
  filter(variable == "B27001_001") # much faster than using distinct()
  select(GEOID, NAME) %>%
  left_join(dat, by = c("GEOID", "NAME")) %>%
  arrange(GEOID) %>%
  rename(uninsured_2012 = `2012`,
    uninsured_2016 = `2016`,
    uninsured_diff = diff)


# Remove the Aleutians West from shp for display purposes.
# NOTE: this isn't necessary since I'm using the shift_geo
# argument in the get_acs function. However if you're not
# using shift_geo or joining to a different spatial layer
# for the full US you may want to consider removing this 
# record for display purposes.
shp <- filter(shp, GEOID != "02016")
```




